---
title: "populate_city_centroid"
output: html_document
---

This RMarkdown document is used to populate the Latitude and Longitude fields in the INWU group's facility list with the city centroid for locations where the coordinates are null. 

This script uses the facilities list "/INWU_task_folders/Site_selection/Industrial_site_list/USEPA_HIFLD_EIA_PPP_facility_v7_NM_NAICS.csv"

and the GNIS layer '\\INWU_task_folders\\Site_selection\\GIS\\Ref_layers\\GNIS\\US_GNIS_for_GIS.csv'

THe output of this script is going to be v7.1 of the facilities list

```{r}
library(tidyverse)
library(sf)

source("../utility_functions/populate_facility_coords_cty_centroid.R")

# Getting User Inputs 

# Path to INWU model folder on local machine
inwu_model_folder <- "D:/DOI/GS-W-WaterUse - Industrial model"

GNIS_path <- paste0(inwu_model_folder, '\\INWU_task_folders\\Site_selection\\GIS\\Ref_layers\\GNIS\\US_GNIS_for_GIS.csv')


```

```{r}

inwu_facility_csv <- paste0(inwu_model_folder,
                            "/INWU_task_folders/Site_selection/Industrial_site_list/USEPA_HIFLD_EIA_PPP_facility_v7_NM_NAICS.csv")

inwu_facility_df <- read_csv(inwu_facility_csv,
                             col_types = cols(
                               LATITUDE = col_double(),
                               LONGITUDE = col_double(),
                               NAICS = col_character(),
                               EMP = col_integer(),
                               TotPerson = col_integer(),
                               JobsReported = col_integer(),
                                                 ))

inwu_facility_df <- populate_facility_coords_cty_centroid(inwu_facility_df, 'LATITUDE', 'LONGITUDE', GNIS_path)

write.csv(inwu_facility_df, '../outputs/USEPA_HIFLD_EIA_PPP_facility_v7.1.csv')
```