# Mapping Coefficient Parameters

## Purpose

This R Markdown Document exists to map out the parameters used for various coefficients.

Coefficients can be found in the team sharepoint at the following location [\~\GS-W-WaterUse - Industrial model\INWU\_task_folders\Lit\_review\Coefficient Values from Literature](https://doimspp.sharepoint.com/:f:/r/sites/USGSWaterUse-Internal/Shared%20Documents/Industrial%20model/INWU_task_folders/Lit_review/Coefficient%20Values%20from%20Literature?csf=1&web=1&e=1AjM1F)

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Importing Requred Packages 

library(tidyverse)
library(ggplot2)
library(sf)
library(viridis)
library(readxl)

source("../utility_functions/download_shape_zip.R")
```

## Getting User Inputs

```{r, include=FALSE}
# Path to INWU model folder on local machine
inwu_model_folder <- "D:/DOI/GS-W-WaterUse - Industrial model/"

export_figures = TRUE
```

## Defining Functions

```{r}

export_figure <- function(figure, path,name, width=10, height=5){
  ggsave(paste0(path, name, ".png"), 
         figure, 
         bg='white', 
         width = width, 
         height = height)
}

```

## Importing Required Data

## Mapping Coefficients By Paper

### Ellis and Others

[Ellis and Others](https://doimspp.sharepoint.com/:b:/r/sites/USGSWaterUse-Internal/Shared%20Documents/Industrial%20model/Literature/SS01_Panel1_Paper03.pdf?csf=1&web=1&e=1bzGiq) has coefficients in gal/ton of product.A table of coefficient values can be found [here](https://doimspp.sharepoint.com/:x:/r/sites/USGSWaterUse-Internal/Shared%20Documents/Industrial%20model/INWU_task_folders/Lit_review/Coefficient%20Values%20from%20Literature/Coeff_IN_WU_Ellis,%20Dillich,%20Margolis,%20n.d.%20(paper%20from%20Energetics,%20Incoporated%20and%20DOE).xlsx?d=w4a94fcc18ee247b79fa03da851c340b8&csf=1&web=1&e=hqLq8u)

These coefficients are nation wide and units are gal/ammount of product (units for ammount vary by product).

Currently we have no data sets for ammountof product.The [Federal Reserve Economic Data](https://fred.stlouisfed.org/series/IPMAN), and [Economic Cenus](https://www.census.gov/programs-surveys/economic-census/data/tables/industry.html) have some production values, but they all appear to be in dollar ammounts.

### Boero and Pasqualini

<https://doi.org/10.1016/j.wri.2017.09.001>

Boero and Pasqualini have National coefficients for the calendar years 1997 - 2013 in the units of gallons of water / GDP at current prices.

To map these data we must downscale national GDP to the county level.

#### Plotting Coefficients

##### Ecoregions

Because these coefficients are mapped to the EPA Level II ecological regions we will need to pull that layer in

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

ecoregions_download_url <- "https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/cec_na/na_cec_eco_l2.zip"

ecoregions_download_path <- '../data/l2_ecoregions.zip'

download_shape_zip(ecoregions_download_url, ecoregions_download_path)

# Clip to CONUS 

us_states_url <- "https://www2.census.gov/geo/tiger/GENZ2018/shp/cb_2018_us_state_500k.zip"

us_states_download_path <- '../data/us_states.zip'

download_shape_zip(us_states_url, us_states_download_path)

# Plot

l2_ecoregions_shp <- '../data/ecoregions/NA_CEC_Eco_Level2.shp'

l2_ecoregions <- read_sf(l2_ecoregions_shp)

plot(l2_ecoregions['NA_L2CODE'])
```

##### Reading coefficients

```{r}

years <- seq(1997, 2013)

boreo_and_pasqualini_coefficients_folder <- paste0(inwu_model_folder, '/INWU_task_folders/Lit_review/Coefficient_Values_from_Literature/Coeff_IN_WU_Boero_Pasqualini')

# Combining into one dataframe
boreo_and_pasqualini_coefficients_df <- data.frame()

for (year in years){ 
  
  coefficient_csv <- paste0(boreo_and_pasqualini_coefficients_folder, '/Coeff_IN_WU_Boero_Pasqualini_', year, '.csv')
  
  coefficient_df <- read.csv(coefficient_csv)
  
  coefficient_df['year'] <- year 
  
  boreo_and_pasqualini_coefficients_df <- rbind(boreo_and_pasqualini_coefficients_df, coefficient_df)
  
  }

# Get a list of region columns to transform dataframe from wide to long
region_columns <- names(boreo_and_pasqualini_coefficients_df)
region_columns <- setdiff(region_columns, c('naics', 'naics_description', 'year'))

# Filtering to only inclue manufacturing naics codes and formatting df as long
boreo_and_pasqualini_coefficients_df <- boreo_and_pasqualini_coefficients_df %>% filter(substr(naics, 1,1) == "3") %>%
  pivot_longer(cols = region_columns,
               names_to = "coefficient",
               values_to = "coefficient_value") %>%
  mutate(l2_ecoregion = gsub("_", ".", sub(".*_(\\d+_\\d+)$", "\\1", coefficient))) %>%
  select(-coefficient)

boreo_and_pasqualini_coefficients_df
```

###### Selecting NAICS Codes with high water useage

many naics have coefficients at or near 0. highest coefficients to see what highest categores are

```{r}

# Summarize mean coefficient value by NAICS and get the top 10
top_naics <- boreo_and_pasqualini_coefficients_df %>%
  group_by(naics, naics_description) %>%
  summarise(mean_coefficient_value = mean(coefficient_value, na.rm = TRUE)) %>%
  arrange(desc(mean_coefficient_value)) %>%
  slice_head(n=10) 

# View the top NAICS
top_naics <- top_naics$naics[0:10]

boreo_and_pasqualini_coefficients_df <- boreo_and_pasqualini_coefficients_df%>%
  filter(naics %in% top_naics)

print(unique(boreo_and_pasqualini_coefficients_df$naics))

```

##### Mapping Coefficients

Creating Maps of coefficient values.
```{r}
naics_list = unique(boreo_and_pasqualini_coefficients_df$naics)

boreo_and_pasqualini_output_dir <-'../outputs/boreo_and_pasqualini_coefficients/'


# Merge to include naics_description
average_coefficients <- boreo_and_pasqualini_coefficients_df %>%
  group_by(naics, l2_ecoregion) %>%
  summarise(average_coefficient = mean(coefficient_value, na.rm = TRUE)) %>%
  left_join(boreo_and_pasqualini_coefficients_df %>% select(naics, naics_description) %>% distinct(), 
            by = "naics") %>%
  ungroup()

for (naics in naics_list) {
  
  # Get the corresponding naics_description
  naics_desc <- average_coefficients %>%
    filter(naics == !!naics) %>%
    pull(naics_description) %>%
    unique()
  
  boreo_and_pasqualini_map <- ggplot(data = merged %>% filter(naics == !!naics)) +
    geom_sf(aes(fill = average_coefficient)) +
    scale_fill_viridis(option = "plasma", na.value = "grey50", trans = "log10") +
    labs(title = paste0("Boreo and Pasqualini coefficients for ", naics, ": ", naics_desc), 
         fill = "gal / $ of GDP") +
    theme_void() +
    theme(legend.position = "right")

  if (export_figures) {
    export_figure(boreo_and_pasqualini_map, boreo_and_pasqualini_output_dir, paste0('map_', naics), 10, 5)
  }
}
```
##### Plotting Coefficients

Here we create plots of the coefficient by ecoregion by year.

```{r}

# Loop through each ecoregion
#for (ecoregion in ecoregion_list) {
for (naics in naics_list) {
    
  # Filter the DataFrame for the desired NAICS codes
  filtered_boreo_and_pasqualini_df <- boreo_and_pasqualini_coefficients_df %>%
    filter(naics == !!naics)  # Adjust this if you want to filter by specific NAICS
  
  # Create the plot, coloring by ecoregion
  boreo_and_pasqualini_plot <- ggplot(filtered_boreo_and_pasqualini_df, aes(x = year, y = coefficient_value, color = l2_ecoregion)) + 
    geom_point() +
    labs(title = "Coefficient Value by Year",
         x = "Year",
         y = "Coefficient Value") + 
    theme_minimal() +
    scale_color_discrete(name = "Ecoregion")  # Optional: Customize the legend title
  
  # Export the figure if the flag is set
  if (export_figures) {
    export_figure(boreo_and_pasqualini_plot, boreo_and_pasqualini_output_dir, 
                  paste0('plot_', naics), 10, 5)
  }
  }
#}
 
 
```

##### Recreating variables

Methodology for estimating county-level GDP from Boreo and Pasqualini:

> “When considering national GDP by sector, we use the data estimated in the IO table just described. When considering county-level GDP by sector, we use local labor market data for its estimation. In particular we use annual employment and average weekly salary from the Quarterly Census of Employment and Wages (QCEW) published by the U.S. Bureau of Labor Statistics at the county level and with three-digit level of detail according to the North American Industrial Classification System (NAICS). QCEW employment and salary data at the county level is then aggregated from the three-digit NAICS level to the BEA summary level. National GDP data by sector is divided by the result of the multiplication between national number of employees and average weekly salary. By doing that, we estimate the average amount of GDP per amount of labor factor employed in each sector and we then estimate county-level GDP by multiplying average national GDP per unit of labor with the amount of labor in the county, similarly to commonly adopted approaches to the regionalization of economic variables (e.g., [29], [3]).”

##### National GDP by Industry

National GDP by Industry is pulled from the Bureau of Economic Analysis using their API.

In order to use this you must [register your email to obtain an API Key](https://apps.bea.gov/API/signup/)

```{r}

bea_api_key = "A0C4BEAD-2048-4D1D-A027-2F47E0D65498"
```

Make a data request for GDP results by industry

```{r}

bea_api_link = paste0("https://apps.bea.gov/api/data?&UserID=", bea_api_key, "&method=GETDATA&DatasetName=GDPbyIndustry&key=10&Year=ALL&Industry=ALL&FREQUENCY=A&TableID=10&ResultFormat=JSON")

bea_api_response <- get(bea_api_link)

bea_content <- content(bea_api_response, 'text')

bea_df <- fromJSON(bea_content, flatten = TRUE)
```

```{r}
# URL of the Excel file
url <- "https://www.bea.gov/sites/default/files/2024-06/gdp1q24-3rd.xlsx"

bea_gdp_path <- '../data/bea_gdp.xlsx'

if (file.exists(bea_gdp_path)) {
  print("bea_gdp.xlsx already exists, using existing data")
} else {
  
  print("Downloading bea_gdp.xlsx")
  download.file(url, bea_gdp_path)
  
}

# Read the Excel file
data <- read_excel(bea_gdp_path)

# View the data
print(data)
```
